### 2023-02-06 appendix.Rmd
revision quality check

---
output:
  pdf_document: default
  html_document: default
---
\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

@@

The minimum size of our sample was determined by conducting a power analysis on data from Peysakhovich et al. (2014). We used the smallest significant correlation between economic game play and real-world cooperation in that study as the effect size (_r_ = 0.15). G*Power (Faul, Erdfelder, Lang, & Buchner, 2007) suggested a sample of 571 participants to detect this correlation effect size with statistical power of 0.95. We aimed to sample 1000 participants, substantially above this amount, and eventually report data from 991 participants.

@@

In Studies 1 and 2, session sizes were often not strict multiples of four necessary for real-time random matching, either due to unforeseen dropouts or uneven sign-ups. To remedy this, "simulated players" were used to round up uneven session sizes. These simulated players chose responses in the games based on the median responses from Peysakhovich et al. (2014). Participants were made aware that they may be matched with simulated players, but only on the payoff screen _after_ all game decisions had been made: "In the rare event that we could not find a participant to match you with, we have instead matched you with average decisions based on previous research."

@@

As with our principal components analysis, we initially restricted our confirmatory factor model to include only the games used in previous work (Peysakhovich et al., 2014). Controlling for game comprehension, we loaded the Dictator Game, Trust Game (Give), Trust Game (Return), and Public Goods Game onto the cooperation latent variable, and loaded the Ultimatum Game (Minimum Acceptable Offer), Third-Party Punishment Game (Punish), and Second-Party Punishment Game (Punish Defector) onto the punishment latent variable. According to established fit statistic cutoffs (Hu & Bentler, 1999; MacCallum, Browne, & Sugawara, 1996), this model fitted the data well, with a Root Mean Square Error of Approximation (RMSEA) value of `r printnum(as.numeric(fitMeasures(cfa1)['rmsea']))`, a Standardised Root Mean Square Residual (SRMR) value of `r printnum(as.numeric(fitMeasures(cfa1)['srmr']))`, and a Comparative Fit Index (CFI) value of `r printnum(as.numeric(fitMeasures(cfa1)['cfi']))`. All indicators had significantly positive loadings, and the two latent variables were significantly negatively correlated with one another (_r_ = `r printnum(standardizedSolution(cfa1)[25,4])`, _p_ = `r printp(parameterEstimates(cfa1)[25,7])`; Supplementary Figure \@ref(fig:cfa1Plot)). Model fit improved further when additionally loading the Stag Hunt Game onto the cooperation latent variable and loading the Stag Hunt Game with Punishment (Punish Defector) onto the punishment latent variable (RMSEA = `r printnum(as.numeric(fitMeasures(cfa2)['rmsea']))`; SRMR = `r printnum(as.numeric(fitMeasures(cfa2)['srmr']))`; CFI = `r printnum(as.numeric(fitMeasures(cfa2)['cfi']))`; Supplementary Figure \@ref(fig:cfa2Plot)).

@@

(ref:beastImagesCaption) _Images used in the BEAST task in rounds 1-5 (a-e)._

@@

(ref:efaTableCaption) Factor loadings from an exploratory factor analysis of all items from the 16-item SDO scale (Sidanius & Pratto, 1999) and the 30-item RWA scale (Altemeyer, 1996). The factor model was fitted to a separate validation dataset of New Zealand participants (n = `r nrow(readd(dVal))`). The twelve items used in the New Zealand Attitudes and Values Study are emphasised in bold.

@@

(ref:testRetestTableCaption) Test-retest reliability for the economic games in Peysakhovich et al. (2014) across Studies 1 and 2. Correlations are Spearman's rank correlations (all _p_ < .050). Intra-class correlations are two-way single score agreements across the two studies (all _p_ < .050).

@@

## Supplementary References

Altemeyer, B. (1996). _The authoritarian specter_. London: Harvard University Press.

Faul, F., Erdfelder, E., Lang, A.-G., & Buchner, A. (2007). G*Power 3: A flexible statistical power analysis program for the social, behavioral, and biomedical sciences. _Behavior Research Methods_, _39_(2), 175–191. https://doi.org/10.3758/BF03193146

Hu, L. T., & Bentler, P. M. (1999). Cutoff criteria for fit indexes in covariance structure analysis: Conventional criteria versus new alternatives. _Structural Equation Modeling_, _6_(1), 1–55. https://doi.org/10.1080/10705519909540118

MacCallum, R. C., Browne, M. W., & Sugawara, H. M. (1996). Power analysis and determination of sample size for covariance structure modeling. _Psychological Methods_, _1_(2), 130–149. https://doi.org/10.1037/1082-989x.1.2.130

Peysakhovich, A., Nowak, M. A., & Rand, D. G. (2014). Humans display a ’cooperative phenotype’ that is domain general and temporally stable. _Nature Communications_, _5_(1), 4939. https://doi.org/10.1038/ncomms5939

Sidanius, J., & Pratto, F. (1999). _Social dominance: An intergroup theory of social hierarchy and oppression_. Cambridge: Cambridge University Press.

>>>>

---
output:
  pdf_document: default
  html_document: default
---
\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
\pagenumbering{gobble}

@@

The minimum size of our sample was determined by conducting a power analysis on data from previous work[@Peysakhovich2014]. We used the smallest significant correlation between economic game play and real-world cooperation in that study as the effect size (_r_ = 0.15). G*Power[@Faul2007] suggested a sample of 571 participants to detect this correlation effect size with statistical power of 0.95. We aimed to sample 1000 participants, substantially above this amount, and eventually report data from 991 participants.

@@

In Studies 1 and 2, session sizes were often not strict multiples of four necessary for real-time random matching, either due to unforeseen dropouts or uneven sign-ups. To remedy this, "simulated players" were used to round up uneven session sizes. These simulated players chose responses in the games based on the median responses from previous work[@Peysakhovich2014]. Participants were made aware that they may be matched with simulated players, but only on the payoff screen _after_ all game decisions had been made: "In the rare event that we could not find a participant to match you with, we have instead matched you with average decisions based on previous research."

@@

As with our principal components analysis, we initially restricted our confirmatory factor model to include only the games used in previous work[@Peysakhovich2014]. Controlling for game comprehension, we loaded the Dictator Game, Trust Game (Give), Trust Game (Return), and Public Goods Game onto the cooperation latent variable, and loaded the Ultimatum Game (Minimum Acceptable Offer), Third-Party Punishment Game (Punish), and Second-Party Punishment Game (Punish Defector) onto the punishment latent variable. According to established fit statistic cutoffs[@Hu1999; @MacCallum1996], this model fitted the data well, with a Root Mean Square Error of Approximation (RMSEA) value of `r printnum(as.numeric(fitMeasures(cfa1)['rmsea']))`, a Standardised Root Mean Square Residual (SRMR) value of `r printnum(as.numeric(fitMeasures(cfa1)['srmr']))`, and a Comparative Fit Index (CFI) value of `r printnum(as.numeric(fitMeasures(cfa1)['cfi']))`. All indicators had significantly positive loadings, and the two latent variables were significantly negatively correlated with one another (_r_ = `r printnum(standardizedSolution(cfa1)[25,4])`, _p_ = `r printp(parameterEstimates(cfa1)[25,7])`; Supplementary Figure \@ref(fig:cfa1Plot)). Model fit improved further when additionally loading the Stag Hunt Game onto the cooperation latent variable and loading the Stag Hunt Game with Punishment (Punish Defector) onto the punishment latent variable (RMSEA = `r printnum(as.numeric(fitMeasures(cfa2)['rmsea']))`; SRMR = `r printnum(as.numeric(fitMeasures(cfa2)['srmr']))`; CFI = `r printnum(as.numeric(fitMeasures(cfa2)['cfi']))`; Supplementary Figure \@ref(fig:cfa2Plot)).

@@

(ref:beastImagesCaption) _Images used in the BEAST task in rounds 1-5 (a-e)._ These images were generated by adapting LIONESS code available at https://github.com/LucasMolleman/Molleman_Kurvers_vandenBos_EHB2019_BEAST

@@

(ref:efaTableCaption) Factor loadings from an exploratory factor analysis of all items from the 16-item SDO scale[@Sidanius1999] and the 30-item RWA scale[@Altemeyer1996]. The factor model was fitted to a separate validation dataset of New Zealand participants (n = `r nrow(readd(dVal))`). The twelve items used in the New Zealand Attitudes and Values Study are emphasised in bold.

@@

(ref:testRetestTableCaption) Test-retest reliability for the economic games from previous work[@Peysakhovich2014] across Studies 1 and 2. Correlations are Spearman's rank correlations (all _p_ < .050). Intra-class correlations are two-way single score agreements across the two studies (all _p_ < .050).

@@

### 2023-02-06 manuscript.Rmd
revision quality check

note              : "This working paper has not yet been peer-reviewed."
keywords          : "cooperation, conformity, punishment, behavioural economics, political ideology"
wordcount         : "5217 words excluding Methods (Methods: 3373 words)"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
  - \renewcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth - 1ex}{#1}\break}
appendix          : "appendix.Rmd"
csl               : "nature.csl"
urlcolor          : blue
notes-after-punctuation: false
---

>>>>

note              : "This working paper has not yet been peer-reviewed."
keywords          : "cooperation, conformity, punishment, behavioural economics, political ideology"
wordcount         : "5217 words excluding Methods (Methods: 3373 words)"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
  - \renewcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth - 1ex}{#1}\break}
appendix          : "appendix.Rmd"
csl               : "nature.csl"
urlcolor          : blue
notes-after-punctuation: false
---
